---
title: "Informatics HW"
author: "Carson"
date: "4/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Impoting the Data!!!

```{r}
setwd("C:/Users/carso/OneDrive/Desktop/Informatics HW")

library(readr)

# Groundwater
gwd <- read.csv("gwd_all_raw.csv", row.names = 1) 


gwd[is.na(gwd)] <- 0



# Physical factors
pred1 <- read.csv("gwd_pred.csv", row.names = 1) 


pred1[is.na(pred1)] <- 0


pred.norm <- decostand(pred1, "normalize")


```

Looking at the Data!?!?!? 
```{r}


raw <- gwd
pred <- pred1

View(raw)
View(pred)

```


### Making tabs!!

## (iii) -PCA{.tabset}
```{r}library("FactoMineR")
library("factoextra")
```
### __Physical & chemical data__



## Analysis 

#### Packages
```{r}
library(ade4)
library(vegan)
library(gclus)
library(ape)

```

Just data - found online 
```{r}
raw.pca1 <- prcomp(raw, center = TRUE,scale. = TRUE)
raw.pca1
```

## Water stuff :)
```{r}

#PCA


raw.pca <- rda(raw, scale = TRUE)
raw.pca

raw1 <- raw.pca$CA$eig

source("evplot.R")
evplot(raw1)
```
Broken stick model:

% Variation explained in the model for each pc - use pc1 and pc2 because the explain the most variance in my model - probably around 70% together!!


# Physical 
```{r}
pred.pca <- rda(pred, scale = TRUE)
pred.pca

pred1 <- pred.pca$CA$eig

source("evplot.R")
evplot(pred1)
```


## Basic clean PCA Water 
```{r}
source("cleanplot.pca.R")
cleanplot.pca(raw.pca, point = TRUE) 
```
 - alkalinity out of the circle etc. 
 - 

### physical 
```{r}
source("cleanplot.pca.R")
cleanplot.pca(pred.pca, point = TRUE) 
```
DO_sat and diss right angle? 


###Basics
Any given pair of
axes, the variables that have vectors longer than this radius
make a higher contribution than average and can be
interpreted with confidence.
ex. Alkalinity 




## Clustering with water quality
```{r}
#Transform
raw.hel <- decostand(raw, "hellinger")

raw.w <- hclust(dist(scale(raw)), "ward")

raw.w

gr <- cutree(raw.w, k = 4)

gr1 <- levels(factor(gr))

raw.sc1 <- scores(raw.pca, display = "wa", scaling =1)

 z <- plot(raw.pca, display = "wa", scaling = 1,
type = "n", main = "PCA correlation with
clusters for ground water data")

 for (i in 1:length(gr1)) {points(raw.sc1[gr==i,],
pch = (14 + i), cex = 2, col = i+1)}

     text(raw.sc1, row.names(raw), cex = 0.7, pos = 3)
    
      ordicluster(z, raw.w, col = "black")
      
     


```

Can see 4 groups of clusters!! - gives us a good idea at what variables are similiar and have the most impact.

## More clustering with physical factors:

```{r}
pred.pca <- rda(pred, scale = TRUE)


 pred.hel <- decostand(pred, "hellinger")

pred.w <- hclust(dist(scale(pred)), "ward")

pred.w

gr3 <- cutree(pred.w, k = 4)

gr4 <- levels(factor(gr3))

pred.sc1 <- scores(pred.pca, display = "wa", scaling =1)

 z <- plot(pred.pca, display = "wa", scaling = 1,
type = "n", main = "PCA correlation with
clusters for physical factors")

 for (i in 1:length(gr4)) {points(pred.sc1[gr3==i,],
pch = (14 + i), cex = 2, col = i+1)}

     text(pred.sc1, row.names(pred), cex = 0.7, pos = 3)
    
      ordicluster(z, pred.w, col = "black")
```


Heatmap!
```{r}
dend <- as.dendrogram(raw.ch.ward)
heatmap(as.matrix(raw.ch), Rowv=dend, symm=TRUE,
margin=c(3,3))

```
“Hot” (white hot) values represent high similarities and are located
close to the diagonal.




## Doesn't work well... but was fun

### NMDS!!!!

```{r}
 raw.nmds <- metaMDS(raw, distance = "bray")

 raw.bray <- vegdist(raw)
 raw.nmds$stress

 plot(raw.nmds, type = "t", main =
paste("NMDS/Bray - Stress=",
round( raw.nmds$stress,3)))
 
 par(mfrow=c(1,2))
stressplot(raw.nmds, main = "Shepard Plot")
gof = goodness(raw.nmds)

 plot(raw.nmds, type = "t", main = "Goodness of
fit")
 points(raw.nmds, display = "sites", cex =
gof*200)

 
```

Not a huge scatter! - pretty well perserved 
Small bubble - good fit!

```{r}
raw.bray.ward <- hclust(raw.bray, "ward")
raw.bw.groups <- cutree(raw.bray.ward, k=4)
grp.lev <- levels(factor(raw.bw.groups))

raw.sc <- scores(raw.nmds)
v <- ordiplot(raw.sc, type = "n", main =
"NMDS/Bray - clusters Ward/Bray")

for (i in 1:length(grp.lev))
{points(raw.sc[raw.bw.groups==i,], pch=(14-i),
cex=2, col=i+1)}

text(raw.sc, row.names(raw), pos=2, cex=0.5)
ordicluster(v, raw.bray.ward, col="black")

ordiellipse(raw.nmds, groups = raw.bw.groups,
draw = "polygon", lty = 1, col = "grey90")

raw.fit <- envfit(raw.nmds, raw, permutations =
999)

plot(raw.fit, p.max = 0.01, col = "black", cex =
0.7)




```

```{r}
pred.nmds <- metaMDS(Pred, distance = "bray")
 pred.bray <- vegdist(Pred)
 pred.bray.ward <- hclust(pred.bray, "ward")
pred.bw.groups <- cutree(pred.bray.ward, k=4)
grp.lev2 <- levels(factor(pred.bw.groups))
sit.sc2 <- scores(pred.nmds)

p2 <- ordiplot(sit.sc2, type = "n", main =
"NMDS/Bray - clusters Ward/Bray")
for (i in 1:length(grp.lev2))
{points(sit.sc2[pred.bw.groups==i,], pch=(3-i),
cex=2, col=i+8)}

text(sit.sc2, row.names(pred), pos=4, cex=0.7)

ordihull(pred.nmds, groups = pred.bw.groups, draw =
"polygon", lty = 1, col = "grey90") 

pred.fit <- envfit(pred.nmds, Pred, permutations =
999)

plot(pred.fit, p.max = 0.01, col = "black", cex =
0.7)

```











